---
title: "Working Title"
author: "__Dr Massimiliano Canzi__"
date: "May 5th 2022"
header-includes:
  - \usepackage{tipa}
  - \usepackage{ulem}
output: 
  beamer_presentation:
    latex_engine: xelatex
    theme: "Frankfurt"
    colortheme: "dove"
    font: "serif"
classoption: "aspectratio=169"
---

```{r}
library(ggpubr)
library(lme4)
library(lmerTest)
library(sjPlot)
library(tidyverse)
library(viridis)
```

```{r setup, include = F}
knitr::opts_chunk$set(fig.align="center", include = F)
```

# Introduction
 
## Who am I

\pause

\begin{itemize}[<+->]
\item Lab Manager / Data Scientist @ Uni Konstanz
\item PhD in Linguistics @ Uni of Manchester 
\item MSc Forensic Speech Science @ Uni of York
\item MA Linguistics @ Uni of Manchester 
\end{itemize}

## Who am I

Interested in R, experimental design, reproducibility, open science
 
## Once Upon a Time... in Hollywood

\pause Planning a (linguistics) experiment and filming a Hollywood movie are not as far apart as you would think: 

\begin{itemize}[<+->]
\item Write a script
\item Set a production budget 
\item Casting, locations, props, storyboards
\item Shoot the film
\item Editing, colour grading, VFX
\item Press, festivals
\end{itemize}

## Once Upon a Time... in Hollywood

\pause Planning a (linguistics) experiment and filming a Hollywood movie are not as far apart as you would think: 

\begin{itemize}[<+->]
\item Formulate a research question
\item Budget, weight methodologies, availability of participants and tools
\item Recruitment, equipment setup, hardware and software
\item Data collection
\item Data wrangling, analaysis and visualisation
\item Paper, conferences
\end{itemize}

## Once Upon a Time... in Hollywood

Just like when shooting a film. Plan ahead. \pause

The more things you can accurately predict and plan, the more solid your design and experiment will be. That's pretty much all there is to it. 

## Thank you! 

Questions?

## Just kidding

\pause

__Important:__ Starting now, most of the advice in this presentation is based on personal experience. As often is the case, there are many ways one can reach a destination. I am simply presenting some of the tips that helped me along the way. 

# Pre-Production

## Research Questions

\pause

Find a link between the theory and the operalisation. 

\pause 

\begin{itemize}[<+->]
\item What method is best to answer the question? 
\item How many experiments? 
\item What experimental design?
\item How are variables going to be coded?
\end{itemize}

## Methods 

\pause

Choosing an experimental method is equivalent to choosing the __resolution__ of your study.

## Experiment Number

\pause

- Helps limit the design of the study 
- Follow-up studies allow for clarification

## Experimental design

\pause

The experimental design is, obviously, directly correlated to the research question and the hypotheses. 

\pause

It is really important that your design allows you to confidently test your hypotheses as you intend to. 

\pause

\begin{itemize}[<+->]
\item Know your limits! 
\item Know your goals! 
\item Choose a balanced design
\item Register your report
\end{itemize}

# Production

# Post-Production

## Pre-processing and Analysis

Let's have a look at some R code

```{r, include = F}
fm <- read_csv("data/facemask.csv") %>% 
  filter(correct == TRUE) %>%
  select(participant_private_id, participant_device_type, item, fricative, condition, rt) %>%
  mutate_if(is.character, as.factor)
```

```{r}
head(fm)
```

```{r}
rt_log <- fm %>% mutate(rt = log(rt))
```

```{r}
model_one <- rt_log %>% lmer(formula = rt ~ condition * fricative + participant_device_type + (condition | participant_private_id) + (1 | item))
```

```{r}
model_two <- fm %>% 
  #filter(rt >= median(rt) - 3 * mad(rt) & rt <= median(rt) + 3 * mad(rt)) %>%
    glmer(formula = rt ~ condition * fricative + participant_device_type + (1 | participant_private_id) + (1 | item), 
                          family = Gamma(link = "log"), 
                          control = glmerControl(optimizer = "bobyqa"))
```

```{r}
model_one <- rt_log %>% 
  filter(rt >= median(rt) - 3 * mad(rt) & rt <= median(rt) + 3 * mad(rt)) %>%
  lmer(formula = rt ~ condition * fricative + participant_device_type + (condition | participant_private_id) + (1 | item))
```

```{r}
model_two <- fm %>% 
  filter(rt >= median(rt) - 3 * mad(rt) & rt <= median(rt) + 3 * mad(rt)) %>%
    glmer(formula = rt ~ condition * fricative + participant_device_type + (1 | participant_private_id) + (1 | item), 
                          family = Gamma(link = "log"), 
                          control = glmerControl(optimizer = "bobyqa"))
```

```{r}
car::Anova(model_one, type = "III")
car::Anova(model_two, type = "III")
```

```{r, include = F}
ggarrange(plot_model(model_one, type = "pred", terms = c("fricative", "condition")),
          plot_model(model_two, type = "pred", terms = c("fricative", "condition")),
          common.legend = TRUE) -> arranged_plot
```

```{r, width = "50%"}
arranged_plot
```

## Pre-processing and Analysis

\pause

Welcome back!

## Pre-processing and Analysis

\pause

Sometimes, it's not as straightforward i.e. the story of event-related potentials (ERP)

## Reporting results

\pause

\begin{itemize}[<+->]
\item Standardisation
\item Supporting material
\item package \texttt{report}
\end{itemize}

## Reporting results

\footnotesize

We fitted a linear mixed model (estimated using REML and nloptwrap optimizer) to predict rt with condition, fricative and participant_device_type (formula: rt ~ condition * fricative + participant_device_type). The model included condition, participant_private_id and item as random effects (formula: list(~condition | participant_private_id, ~1 | item)). The model's total explanatory power is substantial (conditional R2 = 0.41) and the part related to the fixed effects alone (marginal R2) is of 0.08. The model's intercept, corresponding to condition = NM, fricative = FF and participant_device_type = computer, is at 6.89 (95% CI [6.81, 6.97], t(11044) = 171.01, p < .001). Within this model:

  - The effect of condition [WM] is statistically non-significant and positive (beta = 4.10e-03, 95% CI [-0.04, 0.05], t(11044) = 0.17, p = 0.867; Std. beta = 7.57e-03, 95% CI [-0.08, 0.10])
  - The effect of fricative [SH] is statistically significant and negative (beta = -0.22, 95% CI [-0.30, -0.15], t(11044) = -5.52, p < .001; Std. beta = -0.41, 95% CI [-0.56, -0.27]) ...

## Reporting results

\footnotesize

Analyses were conducted using the R Statistical language (version 4.0.3; R Core Team, 2020) on macOS Big Sur 10.16, using the packages ggpubr (version 0.4.0; Alboukadel Kassambara, 2020), Matrix (version 1.3.2; Douglas Bates and Martin Maechler, 2021), lme4 (version 1.1.26; Douglas Bates et al., 2015), ggplot2 (version 3.3.5; Wickham. ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New York, 2016.), stringr (version 1.4.0; Hadley Wickham, 2019), tidyr (version 1.1.2; Hadley Wickham, 2020), forcats (version 0.5.1; Hadley Wickham, 2021), readr (version 1.4.0; Hadley Wickham and Jim Hester, 2020), dplyr (version 1.0.4; Hadley Wickham et al., 2021), tibble (version 3.1.5; Kirill Müller and Hadley Wickham, 2021), lmerTest (version 3.1.3; Kuznetsova A et al., 2017), purrr (version 0.3.4; Lionel Henry and Hadley Wickham, 2020), sjPlot (version 2.8.9; Lüdecke D, 2021), viridis (version 0.5.1; Simon Garnier, 2018), viridisLite (version 0.4.0; Simon Garnier et al., 2021) and tidyverse (version 1.3.0; Wickham et al., 2019).


